---
title: "Peer Assessment 1"
author: "J.S."
date: "14 de Agosto de 2014"
output: html_document
---

This is the report for the first Peer assessment of Reproducible Research.



### Loading and preprocessing the data

This part of the code will verify if you have the necessary data. In case the data does not exist this procedure will download and unzip the file to your current working directory.
```{r}
dirName <- 'repdata_data_activity'
if (!file.exists(dirName)) {
fileUrl <- 'https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip'
fileName <- 'repdata_data_activity.zip'
download.file(fileUrl, fileName, 'curl')
unzip(fileName)
}
```

```{r,echo=TRUE,eval=FALSE}
library(Hmisc)
```

Loading the data into a dataframe and looking at the structure using **str()** and **head()** function.
```{r,echo=TRUE}
data.activity <- read.csv('./activity.csv')
str(data.activity)
head(data.activity)
data.activity$date <- as.Date(data.activity$date, "%Y-%m-%d")
str(data.activity)
```


The variables included in this data set are:

- steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)
- date: The date on which the measurement was taken in YYYY-MM-DD format
- interval: Identifier for the 5-minute interval in which measurement was taken

#### What is the total number of steps taken per day?
Aggregating the data per Date.
```{r,echo=TRUE}
aggregate.data<- aggregate(data.activity$steps, list(date=data.activity$date), sum,na.rm=TRUE)
names(aggregate.data)<- c("date","total")
aggregate.data$date <- as.factor(aggregate.data$date )
```

```{r,echo=TRUE}
library(ggplot2)
library(scales)

```

Plotting the sum of steps per date.
```{r plots,echo=TRUE, fig.height=4,fig.width=12}
ggplot(aggregate.data, aes(x=date, y=total))+geom_bar(stat="identity")+theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
```

Reporting the mean and median per day.
```{r,echo=TRUE}
#Get mean per day 
aggregate(data.activity$steps, list(dates=data.activity$date), mean,na.rm=TRUE)

#Get median per day 
aggregate(data.activity$steps, list(dates=data.activity$date), median,na.rm=TRUE)
```

#### What is the average daily activity pattern?


Calculating of the average number of steps per day.
```{r,echo=TRUE}
prepplot<-aggregate(data.activity$steps, list(dates=data.activity$interval), median,na.rm=TRUE)
names(prepplot)<-c("interval","avsteps")
```

Plotting the average steps per interval and reporting the highest value.
```{r timeseriesplot,echo=TRUE, fig.width=14,fig.height=6}
# This how I obtained the maximum average number of steps per interval
ind<-which(prepplot$avsteps==max(prepplot$avsteps))
prepplot[ind,]

#pplot of the time series + a dot for the maximum value
 ggplot(prepplot, aes(x = interval, y = avsteps),lwd = 1.3, colour = "red") + geom_line(colour = I("blue"))+geom_point(data=prepplot, aes(x = prepplot[ind,]$interval, y = prepplot[ind,]$avsteps, colour = "red"                                                                                                   )) + theme(legend.position="none") +xlab("Interval")  +ylab("Steps") 
```


#### Imputing missing values

Number of missing Values.
```{r,echo=TRUE}

sum(is.na(data.activity$steps))

```

I choosed to input the missing values using the mean value per interval.
```{r,echo=TRUE}
#inputing the values using the average per interval
to.input<-as.data.frame(aggregate(data.activity$steps, list(interval=data.activity$interval), mean,na.rm=TRUE))

#merging the inputed the mean number os steps per interval.
data.activity.input<- merge(x=data.activity,y=to.input,by = "interval")

#New variable steps.input
data.activity.input$steps.imputed<-data.activity.input$steps

#Replacing the NAs with the average per interval
data.activity.input$steps.imputed[is.na(data.activity.input$steps.imputed)]<-data.activity.input$steps.imputed[is.na(data.activity.input$steps.imputed)]

```

```{r,echo=TRUE,fig.width=14,fig.height=6}

aggregate.data.inputed<- aggregate(data.activity.input$steps.imputed, list(date=data.activity.input$date), sum,na.rm=TRUE)
names(aggregate.data.inputed)<- c("date","total")
aggregate.data.inputed$date <- as.factor(aggregate.data.inputed$date)

ggplot(aggregate.data.inputed, aes(x=date, y=total))+geom_bar(stat="identity", alpha = 0.2)+theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

```

Reporting the mean and median per day of the inputed data.

```{r,echo=TRUE}
#Get mean per day 
aggregate(data.activity.input$steps, list(dates=data.activity.input$date), mean,na.rm=TRUE)

#Get median per day 
aggregate(data.activity.input$steps, list(dates=data.activity.input$date), median,na.rm=TRUE)
```



#### Are there differences in activity patterns between weekdays and weekends?

```{r,echo=TRUE}
#Getting the Week Days

data.activity.input$weekday<-weekdays(as.Date(data.activity.input$date,"%Y-%m-%d"))
data.activity.input$weekdaypt<-"Working Days"
# The days of the week are on my language - Portuguese
data.activity.input$weekdaypt[data.activity.input$weekday %in% c("Domingo", "Sábado")] <- "Weekend"
data.activity.input$weekdaypt<-as.factor(data.activity.input$weekdaypt)


```

```{r,echo=TRUE}
prepplot2<-aggregate(data.activity.input$steps.imputed, list(dates=data.activity.input$interval,weekday=data.activity.input$weekdaypt), mean,na.rm=TRUE)
names(prepplot2)<-c("interval","weekday","avgstepsinput")

```

```{r,echo=TRUE,fig.width=12}

 ggplot(prepplot2, aes(x = interval, y = avgstepsinput, group = weekday)) +geom_line(color="blue")+facet_wrap( ~ weekday) +facet_grid(weekday ~ .)  +xlab("Intervals")  +ylab("Average Steps") 
        
```






